---
title: "Using tubern"
author: "Gaurav Sood"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using tuber}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## tubern: Access YouTube Analytics API via R


### Install, Load the package

To install the latest version from CRAN: 

```{r, eval=FALSE, install}
install.packages("tubern")
```

The latest development version of the package will always be on GitHub. Instructions for installing the package from Github are provided below.

```{r, eval=FALSE, install_g}
# install.packages('devtools')
devtools::install_github("soodoku/tubern", build_vignettes = TRUE)
```

Next, load the package: 

```{r, eval=FALSE, loadlib}
library(tubern)
```

### Using the package

To get going, get the application id and password from Google Developer Console (see [https://developers.google.com/youtube/reporting/](https://developers.google.com/youtube/reporting/)). Enable YouTube APIs. Then set the application id and password via the `yt_oauth` function. For more information about YouTube OAuth, see [YouTube OAuth Guide](https://developers.google.com/youtube/v3/guides/authentication).


```{r, eval=FALSE, auth}
yt_oauth("998136489867-5t3tq1g7hbovoj46dreqd6k5kd35ctjn.apps.googleusercontent.com", "MbOSt6cQhhFkwETXKur-L9rN")
```

```{r, eval=FALSE, prints}
## Waiting for authentication in browser...
## Press Esc/Ctrl + C to abort
## Authentication complete.
```

#### Get Statistics of a Video


```{r, eval=FALSE, getstats}
get_report(ids = "channel==MINE", metrics = "views", start_date = "2010-04-01", end_date ="2017-04-01")
```


```
## $kind
## [1] "youtubeAnalytics#resultTable"

## $columnHeaders
## $columnHeaders[[1]]
## $columnHeaders[[1]]$name
## [1] "views"

## $columnHeaders[[1]]$columnType
## [1] "METRIC"

## $columnHeaders[[1]]$dataType
## [1] "INTEGER"



## $rows
## $rows[[1]]
## $rows[[1]][[1]]
## [1] 6

```
