---
title: "tubern: Comprehensive YouTube Analytics Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tubern: Comprehensive YouTube Analytics Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# tubern: Enhanced YouTube Analytics for R

This vignette demonstrates the enhanced features of tubern, showing you how to efficiently retrieve, analyze, and visualize YouTube Analytics data.

## Quick Setup

```{r setup}
library(tubern)

# First-time setup - shows setup guide
yt_oauth()

# Or with your credentials
yt_oauth("your-client-id.apps.googleusercontent.com", "your-client-secret")
```

## New Features Overview

### 1. Relative Date Ranges

No more manual date calculations! Use intuitive relative dates:

```{r dates}
# Get common date range options
get_common_date_ranges()

# Use relative dates in reports
report <- get_report(
  ids = "channel==MINE",
  metrics = "views,likes",
  start_date = "last_30_days"  # End date calculated automatically
)

# More examples
get_report(ids = "channel==MINE", metrics = "views", start_date = "this_month")
get_report(ids = "channel==MINE", metrics = "views", start_date = "last_quarter")
get_report(ids = "channel==MINE", metrics = "views", start_date = "yesterday")
```

### 2. Helper Functions for Common Reports

#### Channel Overview
```{r channel_overview}
# Get comprehensive channel performance
overview <- get_channel_overview("last_30_days")

# Customize what metrics to include
overview <- get_channel_overview(
  "this_month", 
  include_engagement = TRUE,
  include_subscribers = TRUE
)
```

#### Top Performing Videos
```{r top_videos}
# Get top 10 videos by views
top_videos <- get_top_videos("last_7_days")

# Get more videos with custom metrics
top_videos <- get_top_videos(
  "last_30_days",
  max_results = 25,
  metrics = c("views", "likes", "comments", "shares", "estimatedMinutesWatched")
)
```

#### Audience Demographics
```{r demographics}
# Full demographic breakdown
demographics <- get_audience_demographics("last_90_days")

# Age groups only
age_data <- get_audience_demographics("this_month", dimension = "ageGroup")

# Gender breakdown only
gender_data <- get_audience_demographics("this_quarter", dimension = "gender")
```

#### Geographic Performance
```{r geographic}
# Top countries by views
geo_performance <- get_geographic_performance("last_30_days")

# US states/provinces (requires filtering to US)
us_states <- get_geographic_performance(
  "this_month", 
  dimension = "province",
  filters = "country==US"
)
```

#### Device and Platform Analysis
```{r devices}
# Performance by device type
device_performance <- get_device_performance("last_30_days")

# Operating system breakdown
os_performance <- get_device_performance("this_month", dimension = "operatingSystem")

# Both device and OS
device_os <- get_device_performance(
  "last_7_days", 
  dimension = c("deviceType", "operatingSystem")
)
```

#### Daily Time Series Data
```{r daily}
# Daily performance trends
daily_data <- get_daily_performance("last_30_days")

# With engagement metrics
daily_engagement <- get_daily_performance(
  "this_month",
  metrics = c("views", "likes", "comments", "shares")
)
```

#### Revenue Reports (Requires Monetary Scope)
```{r revenue}
# Set up OAuth with monetary scope first
yt_oauth("your-client-id", "your-client-secret", scope = "monetary")

# Get revenue data
revenue <- get_revenue_report("last_month")

# Revenue in different currency
revenue_eur <- get_revenue_report("this_quarter", currency = "EUR")
```

### 3. Enhanced Parameter Validation

tubern now provides helpful validation and suggestions:

```{r validation}
# Invalid metrics get helpful suggestions
# get_report(ids = "channel==MINE", metrics = "vews", start_date = "last_7_days")
# Error: Invalid metric(s): vews
# Did you mean: 'vews' -> 'views'

# Check available metrics and dimensions
get_available_metrics()
get_available_metrics("view")  # Filter by pattern

get_available_dimensions() 
get_available_dimensions("country|city")  # Geographic dimensions
```

### 4. Data Transformation and Analysis

#### Convert to Data Frames
```{r dataframes}
# Get data and convert to clean data.frame
report <- get_daily_performance("last_30_days")
df <- yt_to_dataframe(report)
head(df)

# Convert to tibble (if tibble package installed)
tbl <- yt_to_tibble(report)

# Keep original column names
df_orig <- yt_to_dataframe(report, clean_names = FALSE)
```

#### Data Export
```{r export}
# Export to CSV
report <- get_top_videos("last_7_days")
file_path <- yt_export_csv(report, "top_videos.csv")

# Auto-generated filename with timestamp
file_path <- yt_export_csv(report)
```

#### Quick Visualization
```{r plots}
# Automatic plot creation (requires ggplot2)
daily_report <- get_daily_performance("last_30_days")
yt_quick_plot(daily_report)  # Line plot over time

# Bar chart for top videos
top_videos <- get_top_videos("last_7_days")
yt_quick_plot(top_videos, chart_type = "bar")

# Custom x and y columns
geo_data <- get_geographic_performance("last_30_days")
yt_quick_plot(geo_data, x_col = "country", y_col = "views")
```

#### Data Summary
```{r summary}
# Get summary statistics
report <- get_channel_overview("last_30_days")
summary <- yt_extract_summary(report)
print(summary)
```

### 5. Improved Error Handling and Diagnostics

#### Diagnostic Tools
```{r diagnostics}
# Check if everything is set up correctly
diagnose_tubern()

# Check API quota status
check_api_quota()
```

#### Error Recovery
```{r errors}
# tubern now provides helpful error messages and suggestions
# for authentication, API quota, parameter errors, etc.

# Example: If OAuth token expires, tubern will suggest running yt_oauth() again
# Example: If API quota is exceeded, tubern will suggest ways to reduce usage
```

## Complete Workflow Examples

### Example 1: Channel Performance Dashboard
```{r workflow1}
# Set up authentication
yt_oauth("your-client-id", "your-client-secret")

# Get overview data
overview <- get_channel_overview("last_30_days")
overview_df <- yt_to_dataframe(overview)

# Get daily trends
daily <- get_daily_performance("last_30_days")
daily_df <- yt_to_dataframe(daily)

# Get top videos
top_videos <- get_top_videos("last_30_days", max_results = 10)
videos_df <- yt_to_dataframe(top_videos)

# Get geographic performance
geo <- get_geographic_performance("last_30_days")
geo_df <- yt_to_dataframe(geo)

# Create visualizations
yt_quick_plot(daily)      # Daily trend
yt_quick_plot(top_videos) # Top videos
yt_quick_plot(geo)        # Geographic performance

# Export data
yt_export_csv(overview, "channel_overview.csv")
yt_export_csv(daily, "daily_performance.csv")
yt_export_csv(top_videos, "top_videos.csv")
```

### Example 2: Audience Analysis
```{r workflow2}
# Demographic analysis
demographics <- get_audience_demographics("last_90_days")
demo_df <- yt_to_dataframe(demographics)

# Device analysis  
devices <- get_device_performance("last_90_days")
device_df <- yt_to_dataframe(devices)

# Geographic analysis
geography <- get_geographic_performance("last_90_days", max_results = 50)
geo_df <- yt_to_dataframe(geography)

# Combined analysis with dplyr (if available)
if (require(dplyr)) {
  # Top countries by watch time
  top_countries <- geo_df %>%
    arrange(desc(estimated_minutes_watched)) %>%
    head(10)
  
  print(top_countries)
}
```

### Example 3: Content Performance Analysis
```{r workflow3}
# Analyze video performance over time
videos_monthly <- get_top_videos("this_month", max_results = 50)
videos_df <- yt_to_dataframe(videos_monthly)

# Compare with previous month
videos_prev <- get_top_videos("last_month", max_results = 50) 
videos_prev_df <- yt_to_dataframe(videos_prev)

# Daily performance for trend analysis
daily_trends <- get_daily_performance(
  "last_30_days",
  metrics = c("views", "estimatedMinutesWatched", "likes", "comments")
)
trends_df <- yt_to_dataframe(daily_trends)

# Visualize trends
yt_quick_plot(daily_trends, x_col = "day", y_col = "views")
```

## Best Practices

### 1. Authentication
- Use `scope = "monetary"` only when you need revenue data
- Keep your OAuth credentials secure
- Use `diagnose_tubern()` to troubleshoot authentication issues

### 2. Data Retrieval
- Use relative dates for convenience: `"last_30_days"`, `"this_month"`
- Start with helper functions like `get_channel_overview()` before using `get_report()` directly
- Use `max_results` parameter to limit data when testing

### 3. Error Handling
- Always check `diagnose_tubern()` if you encounter issues
- Use `check_api_quota()` to monitor your API usage
- Enable `setup_guide = TRUE` for first-time setup help

### 4. Data Analysis
- Convert API responses to data.frames with `yt_to_dataframe()`
- Use `yt_quick_plot()` for rapid data exploration
- Export data with `yt_export_csv()` for external analysis

## Troubleshooting

### Common Issues

1. **Authentication Problems**
   ```{r}
   # Run diagnostics
   diagnose_tubern()
   
   # Re-authenticate
   yt_oauth("your-client-id", "your-client-secret")
   ```

2. **API Quota Exceeded**
   ```{r}
   # Check quota status
   check_api_quota()
   
   # Reduce data scope
   get_report(ids = "channel==MINE", metrics = "views", 
             start_date = "yesterday", max_results = 10)
   ```

3. **Invalid Parameters**
   ```{r}
   # Check available options
   get_available_metrics()
   get_available_dimensions()
   
   # Use helper functions for validation
   get_channel_overview("last_7_days")  # Pre-validated parameters
   ```

## More Resources

- [YouTube Analytics API Documentation](https://developers.google.com/youtube/analytics)
- [Google Cloud Console](https://console.cloud.google.com) for API setup
- [tubern GitHub Repository](https://github.com/soodoku/tubern) for issues and updates